### 2021-04-22 paper/ase21/analyzer.tex
[ASE21] Updated Figure 2 and related parts

The subtype relation $\subtype \subseteq \tyset \times \tyset$ between types is
depicted in Figure~\ref{fig:subtype}; a directed edge from $\ty'$ to $\ty$
denotes $\ty' \subtype \ty$, and the relation is reflexive and transitive.  The
subtype relation depends on the nominal types defined in ECMAScript.  We
extract the subtype relation for AST types from the JavaScript syntax.
For example, consider the syntax-directed abstract
algorithm at the lower-right part in Figure~\ref{fig:example}.
Because the nonterminal \textit{BindingElement} is the unique alternative of the
\textit{FormalParameter} production, we automatically extract the subtype
relation: \textit{BindingElement} $\subtype$ \textit{FormalParameter}. Using
the subtype relation, the expression $\expr: \ty$ checks whether the
evaluation result of $\expr$ has type $\ty'$ satisfying $\ty' \subtype \ty$.

@@

        \emp
        & \text{otherwise}\\
      \end{array}
    \right.
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
where $\refine$ prunes out infeasible parts in abstract
environments using their conditions as we describe in Section~\ref{sec:refine}.
Then, we define the abstract
semantics $\asem{\prog}$ of a program $\prog$
as the least fixpoint of the abstract transfer function $\atransfer: \adom \rightarrow
\adom$:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}l}
      \asem{\prog} &=& \lim_{n \rightarrow \infty}(\atransfer)^n(\iaelem)\\
      \atransfer(\aelem) &=& \aelem \join \left(
        \bigjoin_{(\lab, \tys) \in \domain{\rmap}}
        \asemi{\getinst(\lab)}(\lab, \tys)(\aelem)
      \right)\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
where $\aelem = (\rmap, \_)$ and $\iaelem$ denotes the initial abstract state.
As described in Section~\ref{sec:overview}, $\iaelem$ contains the entry points of
all syntax-directed algorithms without additional parameters and built-in
algorithms with appropriate abstract environments.
For a syntax-directed algorithm, we construct its abstract environment containing the variable
$\this$ with its production type and other variables for nonterminals.
For example, the upper-left syntax-directed algorithm in Figure~\ref{fig:example} is
initialized with the following abstract environment:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}l}
      \this &\mapsto& \{ \code{AssignmentExpression} \},\\
      \code{LeftHandSideExpression} &\mapsto& \{ \code{LeftHandSideExpression} \},\\
      \code{AssignmentExpression} &\mapsto& \{ \code{AssignmentExpression} \}
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
For built-in algorithms, we assign pre-defined variables $\this$, $\args$, and
$\NewTarget$ with their corresponding types and parameters with $\tjs$ types.  For
example, the following abstract environment is for the built-in
algorithm $\jscode{Math.round}$ at the upper-right part in Figure~\ref{fig:example}:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}lr@{~}c@{~}l}
      \this &\mapsto& \{ \tjs \}, &
      \args &\mapsto& \{ \clist{\tjs} \},\\
      \NewTarget &\mapsto& \{ \code{Object}, \undefval \}, &
      \x &\mapsto& \{ \tjs \}\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure}

>>>>

The subtype relation $\subtype \subseteq \tyset \times \tyset$ between types is
depicted in Figure~\ref{fig:subtype}; a directed edge from $\ty'$ to $\ty$
denotes $\ty' \subtype \ty$, and the relation is reflexive and transitive.  The
subtype relation depends on the nominal types defined in ECMAScript.  We
extract the subtype relation for AST types from the JavaScript syntax.
For example, consider the syntax-directed abstract algorithm in Figure~\ref{fig:example}(c).
Because the nonterminal \textit{BindingElement} is the unique alternative of the
\textit{FormalParameter} production, we automatically extract the subtype
relation: \textit{BindingElement} $\subtype$ \textit{FormalParameter}. Using
the subtype relation, the expression $\expr: \ty$ checks whether the
evaluation result of $\expr$ has type $\ty'$ satisfying $\ty' \subtype \ty$.

@@

        \emp
        & \text{otherwise}\\
      \end{array}
    \right.
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
where $\refine$ prunes out infeasible parts in abstract
environments using their conditions as we describe in Section~\ref{sec:refine}.
Then, we define the abstract
semantics $\asem{\prog}$ of a program $\prog$
as the least fixpoint of the abstract transfer function $\atransfer: \adom \rightarrow
\adom$:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}l}
      \asem{\prog} &=& \lim_{n \rightarrow \infty}(\atransfer)^n(\iaelem)\\
      \atransfer(\aelem) &=& \aelem \join \left(
        \bigjoin_{(\lab, \tys) \in \domain{\rmap}}
        \asemi{\getinst(\lab)}(\lab, \tys)(\aelem)
      \right)\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
where $\aelem = (\rmap, \_)$ and $\iaelem$ denotes the initial abstract state.
As described in Section~\ref{sec:overview}, $\iaelem$ contains the entry points of
all syntax-directed algorithms without additional parameters and built-in
algorithms with appropriate abstract environments.
For a syntax-directed algorithm, we construct its abstract environment containing the variable
$\this$ with its production type and other variables for nonterminals.
For example, the syntax-directed algorithm in Figure~\ref{fig:example}(a) is
initialized with the following abstract environment:
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}l}
      \this &\mapsto& \{ \code{AssignmentExpression} \},\\
      \code{LeftHandSideExpression} &\mapsto& \{ \code{LeftHandSideExpression} \},\\
      \code{AssignmentExpression} &\mapsto& \{ \code{AssignmentExpression} \}
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
For built-in algorithms, we assign pre-defined variables $\this$, $\args$, and
$\NewTarget$ with their corresponding types and parameters with $\tjs$ types.  For
example, the following abstract environment is for the built-in
algorithm $\jscode{Math.round}$ in Figure~\ref{fig:example}(d):
\begin{figure}[H]
  \centering
  \vspace*{-0.5em}
  \resizebox{0.9\columnwidth}{!}{$
    \begin{array}{r@{~}c@{~}lr@{~}c@{~}l}
      \this &\mapsto& \{ \tjs \}, &
      \args &\mapsto& \{ \clist{\tjs} \},\\
      \NewTarget &\mapsto& \{ \code{Object}, \undefval \}, &
      \x &\mapsto& \{ \tjs \}\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure}

### 2021-04-22 paper/ase21/checker.tex
[ASE21] Updated Figure 2 and related parts

      \asemi{\kwlet \; \x = \expr}(\lab, \tys)(\aelem) &=& \left\{
        \begin{array}{ll}
          \text{already defined variable} \; $\x$
          & \text{if} \; \aty = \{ \true \}\\
          \cdots
          & \text{otherwise}\\
        \end{array}
      \right.\\ &&
      \text{where} \; \aty = \aseme{\x \kwexists}(\aelem(\lab, \tys))\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
If the abstract semantics of a variable lookup for $\x$ is a singleton $\{ \tabsent \}$,
$\x$ is always an unknown variable.  For example, consider the syntax-directed algorithm
at the upper-left part in Figure~\ref{fig:example}.
Since the \textbf{GetReferencedName} algorithm is removed, the variable
\code{GetReferencedName} does not exist in abstract environments and its lookup
returns $\{ \tabsent \}$. Thus, the reference checker reports the unknown variable
bug for \code{GetReferencedName}.  For duplicated variable declarations,
the reference checker utilizes the abstract semantics of the existence check
$\aseme{\x \kwexists}$ to see whether the variable $\x$ of each variable declaration is
already defined.

@@

          \cdots &
          \text{otherwise}\\
        \end{array}
      \right.\\
      \qquad \text{where} \; \func = \kwfunc{\f}{\p_1, \cdots, \p_{n_\func},
      \kwsl \cdots, p_{m_\func} \kwsr}{\lab} \;\wedge\;\\
      \qquad \phantom{\text{where}} \; \aty = \aseme{\expr}(\aelem(\lab,
      \tys)).\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
For each function $\func$ in the abstract semantics of the function expression
$\expr$, the arity checker compares the number of arguments with the arity of
$\func$ to detect missing parameters.  For example, consider the lower-left
syntax-directed algorithm in Figure~\ref{fig:example}.  The algorithm invocation
on line 2 is compiled to the following function call instruction:
\[
  \fcode{x = (formals.IteratorBindingInitialization formals)}
\]
using a temporary variable $\x$.  Because it passes only a single argument
$\code{formals}$ even though the function arity is $[3, 3]$, the arity checker reports
missing parameter bugs for two additional parameters $\code{iteratorRecord}$ and
$\code{environment}$.

@@

          \cdots &
          \text{otherwise}\\
        \end{array}
      \right.\\ &
      \text{where} \; \aty = \aseme{\expr}(\aelem(\lab, \tys))\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
It checks whether the abstract semantics of the condition expression $\expr$ 
subsumes $\{ \true \}$.  For example, consider the lower-right syntax-directed
algorithm in Figure~\ref{fig:example}.  The parameter $\code{environment}$ of
this algorithm has an environment record or $\undefval$.
Since type sensitivity divides the abstract types of arguments to
upcasted single types, there are two different abstract environments whose
variable $\code{environment}$ points to $\{ \code{Environment} \}$ or $\{ \undefval \}$.
When $\code{environment}$ is $\{ \code{Environment} \}$, the
abstract semantics of the assertion condition $\code{environment = originalEnv}$ is $\{ \tbool \}$.
Even though we know that the type of $\code{originalEnv}$ is
also $\{ \code{Environment} \}$, because \code{Environment} is not a singleton type,
we cannot conclude that they are the exactly same environment.
Thus, the assertion checker does not report any bug for this case.
However, if $\code{environment}$ is $\{ \undefval \}$, the abstract semantics of the
condition $\code{environment = originalEnv}$ is $\{ \false \}$ because an
environment is never equal to $\undefval$.
Therefore, the assertion checker reports an assertion failure for
the condition $\code{environment = originalEnv}$.

@@

For an example of non-numeric operand bugs, consider the upper-right built-in
algorithm $\code{Math.round}$ in Figure~\ref{fig:example}.  The types of $\x$
and $\n$ are $\{ \tjs \}$ and $\{ \tnum \}$, respectively, because the
\textbf{ToNumber} algorithm always returns number values or abrupt completions,
and the prefix \textbf{?} removes the latter case.  The built-in algorithm
misuses $\x$ rather than $\n$ on lines 3 and 4, and because the expected
abstract type $\{ \tnum, \tbigint \}$ does not subsume $\{ \tjs \}$, the operand
checker reports non-numeric operand bugs.

>>>>

      \asemi{\kwlet \; \x = \expr}(\lab, \tys)(\aelem) &=& \left\{
        \begin{array}{ll}
          \text{already defined variable} \; $\x$
          & \text{if} \; \aty = \{ \true \}\\
          \cdots
          & \text{otherwise}\\
        \end{array}
      \right.\\ &&
      \text{where} \; \aty = \aseme{\x \kwexists}(\aelem(\lab, \tys))\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
If the abstract semantics of a variable lookup for $\x$ is a singleton $\{ \tabsent \}$,
$\x$ is always an unknown variable.  For example, consider the syntax-directed algorithm
in Figure~\ref{fig:example}(a).
Since the \textbf{GetReferencedName} algorithm is removed, the variable
\code{GetReferencedName} does not exist in abstract environments and its lookup
returns $\{ \tabsent \}$. Thus, the reference checker reports the unknown variable
bug for \code{GetReferencedName}.  For duplicated variable declarations,
the reference checker utilizes the abstract semantics of the existence check
$\aseme{\x \kwexists}$ to see whether the variable $\x$ of each variable declaration is
already defined.

@@

          \cdots &
          \text{otherwise}\\
        \end{array}
      \right.\\
      \qquad \text{where} \; \func = \kwfunc{\f}{\p_1, \cdots, \p_{n_\func},
      \kwsl \cdots, p_{m_\func} \kwsr}{\lab} \;\wedge\;\\
      \qquad \phantom{\text{where}} \; \aty = \aseme{\expr}(\aelem(\lab,
      \tys)).\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
For each function $\func$ in the abstract semantics of the function expression
$\expr$, the arity checker compares the number of arguments with the arity of
$\func$ to detect missing parameters.  For example, consider the
syntax-directed algorithm in Figure~\ref{fig:example}(b).  The algorithm invocation
on line 2 is compiled to the following function call instruction:
\[
  \fcode{x = (formals.IteratorBindingInitialization formals)}
\]
using a temporary variable $\x$.  Because it passes only a single argument
$\code{formals}$ even though the function arity is $[3, 3]$, the arity checker reports
missing parameter bugs for two additional parameters $\code{iteratorRecord}$ and
$\code{environment}$.

@@

          \cdots &
          \text{otherwise}\\
        \end{array}
      \right.\\ &
      \text{where} \; \aty = \aseme{\expr}(\aelem(\lab, \tys))\\
    \end{array}
  $}
  \vspace*{-0.5em}
\end{figure} \noindent
It checks whether the abstract semantics of the condition expression $\expr$ 
subsumes $\{ \true \}$.  For example, consider the syntax-directed
algorithm in Figure~\ref{fig:example}(c).  The parameter $\code{environment}$ of
this algorithm has an environment record or $\undefval$.
Since type sensitivity divides the abstract types of arguments to
upcasted single types, there are two different abstract environments whose
variable $\code{environment}$ points to $\{ \code{Environment} \}$ or $\{ \undefval \}$.
When $\code{environment}$ is $\{ \code{Environment} \}$, the
abstract semantics of the assertion condition $\code{environment = originalEnv}$ is $\{ \tbool \}$.
Even though we know that the type of $\code{originalEnv}$ is
also $\{ \code{Environment} \}$, because \code{Environment} is not a singleton type,
we cannot conclude that they are the exactly same environment.
Thus, the assertion checker does not report any bug for this case.
However, if $\code{environment}$ is $\{ \undefval \}$, the abstract semantics of the
condition $\code{environment = originalEnv}$ is $\{ \false \}$ because an
environment is never equal to $\undefval$.
Therefore, the assertion checker reports an assertion failure for
the condition $\code{environment = originalEnv}$.

@@

For an example of non-numeric operand bugs, consider the built-in
algorithm $\code{Math.round}$ in Figure~\ref{fig:example}(d).  The types of $\x$
and $\n$ are $\{ \tjs \}$ and $\{ \tnum \}$, respectively, because the
\textbf{ToNumber} algorithm always returns number values or abrupt completions,
and the prefix \textbf{?} removes the latter case.  The built-in algorithm
misuses $\x$ rather than $\n$ on lines 3 and 4, and because the expected
abstract type $\{ \tnum, \tbigint \}$ does not subsume $\{ \tjs \}$, the operand
checker reports non-numeric operand bugs.

### 2021-04-22 paper/ase21/overview.tex
[ASE21] Updated Figure 2 and related parts

\paragraph{Reference Checker} The example JavaScript program first defines a
variable \jscode{f} without initialization, which has the value \jscode{undefined}.
It then assigns an anonymous function to \jscode{f} using the operator \jscode{??=}.
While the corresponding \textbf{Evaluation} algorithm originally
used the \textbf{GetReferencedName} algorithm to get a reference name on line 4.a,
a contributor removed the \textbf{GetReferencedName} algorithm and replaced all its
invocations with accesses of the field [[ReferencedName]] on October 28, 2020.
However, the contributor missed several cases including the semantics of
\jscode{??=}, which was fixed by another contributor on November 3, 2020.
Thus, the unknown variable bug for \textbf{GetReferencedName} existed for 7 days,
which the reference checker can detect.

\paragraph{Arity Checker} The program finally calls \jscode{f} with an argument \jscode{true}.
During the initialization of the function call, the \textbf{IteratorBindingInitialization}
algorithm is executed with additional parameters \textit{iteratorRecord} and
\textit{environment} to assign argument values to parameters.
However, a contributor missed passing additional arguments to them on line 2 in
\textbf{IteratorBindingInitialization} of \textit{ArrowParameters} on September 6, 2018.
It caused an arity mismatch bug, which existed for 533 days until another
contributor fixed it on February 20, 2020. The arity checker can detect such arity mismatches.

\paragraph{Assertion Checker} During the initialization of the function call,
another specification bug existed in \textbf{IteratorBindingInitialization} of
\textit{FormalParameter}.  The additional \textit{environment} parameter may
contain \jscode{undefined} or an environment record exactly the same as
\textit{originalEnv} as described in the assertion on line 5.
However, the assertion missed that the initial commit
of the open development process on September 22, 2015 introduced
\jscode{undefined}, which led to the assertion failure bug last for 1,297 days until April 10, 2019.
The assertion checker can detect such assertion failures.

\paragraph{Operand Checker} After the function call initialization, the
parameter \jscode{x} has the value \jscode{true}, and \jscode{Math.round} is invoked
with the argument \jscode{true}.  The \jscode{Math.round} built-in library first
converts the given parameter \textit{x} to its corresponding number value
\textit{n} using \textbf{ToNumber}, and then performs the remaining
steps using \textit{n}.  However, a contributor mistakenly used
\textit{x} instead of \textit{n} on lines 3 and 4 on September 11, 2020.
This bug caused the algorithm to compare the boolean value \jscode{true} with the
numeric value 0.5 or 0 in the example JavaScript program.
This bug lived for two days until another contributor fixed it, and
the operand checker can detect such bugs.

>>>>

\paragraph{Reference Checker} The example JavaScript program first defines a
variable \jscode{f} without initialization, which has the value \jscode{undefined}.
It then assigns an anonymous function to \jscode{f} using the operator \jscode{??=}.
While the corresponding \textbf{Evaluation} algorithm in Figure~\ref{fig:example}(a) originally
used the \textbf{GetReferencedName} algorithm to get a reference name on line 4.a,
a contributor removed the \textbf{GetReferencedName} algorithm and replaced all its
invocations with accesses of the field [[ReferencedName]] on October 28, 2020.
However, the contributor missed several cases including the semantics of
\jscode{??=}, which was fixed by another contributor on November 3, 2020.
Thus, the unknown variable bug for \textbf{GetReferencedName} existed for 7 days,
which the reference checker can detect.

\paragraph{Arity Checker} The program finally calls \jscode{f} with an argument \jscode{true}.
During the initialization of the function call, the \textbf{IteratorBindingInitialization}
algorithm in Figure~\ref{fig:example}(b) is executed with additional parameters \textit{iteratorRecord} and
\textit{environment} to assign argument values to parameters.
However, a contributor missed passing additional arguments to them on line 2 in
\textbf{IteratorBindingInitialization} of \textit{ArrowParameters} on September 6, 2018.
It caused an arity mismatch bug, which existed for 533 days until another
contributor fixed it on February 20, 2020. The arity checker can detect such arity mismatches.

\paragraph{Assertion Checker} During the initialization of the function call,
another specification bug existed in \textbf{IteratorBindingInitialization} of
\textit{FormalParameter} in Figure~\ref{fig:example}(c).  The additional \textit{environment} parameter may
contain \jscode{undefined} or an environment record exactly the same as
\textit{originalEnv} as described in the assertion on line 5.
However, the assertion missed that the initial commit
of the open development process on September 22, 2015 introduced
\jscode{undefined}, which led to the assertion failure bug last for 1,297 days until April 10, 2019.
The assertion checker can detect such assertion failures.

\paragraph{Operand Checker} After the function call initialization, the
parameter \jscode{x} has the value \jscode{true}, and \jscode{Math.round} in Figure~\ref{fig:example}(d) is invoked
with the argument \jscode{true}.  The \jscode{Math.round} built-in library first
converts the given parameter \textit{x} to its corresponding number value
\textit{n} using \textbf{ToNumber}, and then performs the remaining
steps using \textit{n}.  However, a contributor mistakenly used
\textit{x} instead of \textit{n} on lines 3 and 4 on September 11, 2020.
This bug caused the algorithm to compare the boolean value \jscode{true} with the
numeric value 0.5 or 0 in the example JavaScript program.
This bug lived for two days until another contributor fixed it, and
the operand checker can detect such bugs.
